What We Eval in the Shadows

### 2021-04-14 paper/main.tex
Explanations about tidyselect

\documentclass[review,nonacm,screen,acmsmall,anonymous=true]{acmart}
\settopmatter{printfolios=true,printccs=true,printacmref=true}

@@

\vspace{2mm}\noindent\emph{Is Past Prologue?} Previous work investigated how
\eval is used in web programming, specifically in websites that use
JavaScript~\cite{pldi10a}. In 2010, 17 of the largest website used the feature.
In 2011, 82\% of the 10,000 most accessed sites used \eval~\cite{ecoop11}. Yet,
the strings passed to \eval, and their behaviors when executed, are far from
random; it was shown that when one can observe several calls to \eval, the
``shape'' of future calls can be predicted with 97\% accuracy~\cite{oopsla12b}.
Overall, practical usage suggested that most reflective calls were relatively
harmless. While this backs up the soundiness squad's approach, does it
generalize to other application domains than web programming and to other
languages?

@@

\subsection{The arguments of \eval}

The expressions passed to \eval vary widely. In order to categorize them, let us
use a minimization function $min(e)$ which for a given expression $e$ returns a
normal form that abstracts incidental details allowing the reader to focus on
the structure of the evaluated code. The minimization function performs constant
folding of arithmetic and string expressions for base operators, e.g.
$min(\c{1+1})=\c{V}$, value simplification, $min(\c{c(1,2,3+2)})=\c{V}$,
variable absorption, $min(\c{x+y})=\c{X}$, function absorption,
$min(\c{g(f(x),h(z))})=\c{F(F(X))}$ and a number of other simplifications.
Table~\ref{tab:minimizedexpressions} gives the 10 most frequent forms; \#sites
and \%sites are, respectively, the number and ratio of sites receiving arguments
of that form; \#packages is the number of packages with that form; \#operations
is the median number of instructions performed by the interpreter; and \%envir
is the ratio of sites for that evaluate in a function environment. The example
column shows one sample expression $e$ that normalizes to the particular form.

@@

The R language was intended to be extensible. The combination of lazy
evaluation, \c{substitute} and \c{eval} are the tools given to
developers to this end. This API is slightly more complex than just
passing a string, it is conceivable that this may discourage some
casual users. \Eval is also being used to reduce boilerplate code and
provide convenience features for programmers. We now give some
representative examples of its usage.

\mypara{Variable lookup.} % predominant usage of eval in base

@@

>>>>

\documentclass[review,screen,acmsmall,anonymous=true]{acmart} 
%\documentclass[noacm,review,screen,acmsmall,anonymous=true]{acmart} 
%Option noacm above messes with the line numbers in review mode
\settopmatter{printfolios=true,printccs=true,printacmref=true}

@@

\vspace{2mm}\noindent\emph{Is Past Prologue?} Previous work investigated how
\eval is used in web programming, specifically in websites that use
JavaScript~\cite{pldi10a}. In 2010, 17 of the largest websites used the feature.
In 2011, 82\% of the 10,000 most accessed sites used \eval~\cite{ecoop11}. Yet,
the strings passed to \eval, and their behaviors when executed, are far from
random; it was shown that when one can observe several calls to \eval, the
``shape'' of future calls can be predicted with 97\% accuracy~\cite{oopsla12b}.
Overall, practical usage suggested that most reflective calls were relatively
harmless. While this backs up the soundiness squad's approach, does it
generalize to other application domains than web programming and to other
languages?

@@

\subsection{The expression in \eval} \label{sec:minimized}

The expressions passed to \eval vary widely. In order to categorize them, let us
use a minimization function $min(e)$ which for a given expression $e$ returns a
normal form that abstracts incidental details allowing the reader to focus on
the structure of the evaluated code. The minimization function performs constant
folding of arithmetic and string expressions for base operators, e.g.
$min(\c{1+1})=\c{V}$, value simplification, $min(\c{c(1,2,3+2)})=\c{V}$,
variable absorption, $min(\c{x+y})=\c{X}$, function absorption,
$min(\c{g(f(x),h(z))})=\c{F(F(X))}$ and a number of other simplifications.
Table~\ref{tab:minimizedexpressions} gives the 10 most frequent forms; \#sites
and \%sites are, respectively, the number and ratio of sites receiving arguments
of that form; \#packages is the number of packages with that form; \#operations
is the median number of instructions performed by the interpreter; and \%envir
is the ratio of sites that evaluate in a function environment. The example
column shows one sample expression $e$ that normalizes to the particular form.

@@

The R language was intended to be extensible. The combination of lazy
evaluation, \c{substitute} and \c{eval} are the tools given to
developers to this end. This API is slightly more complex than just
passing a string, it is conceivable that this may discourage some
casual users. \Eval is also being used to reduce boilerplate code and
provide convenience features for programmers. We give examples for the minimized expressions of Section~\ref{sec:minimized} and then dig into high-level \eval design patterns.

\mypara{Variable lookup and values.} 
This corresponds to minimized expressions \c{X}, \c{V}. We also include here lookup to slots in a list, \ie minimized expression \c{\$}.
It is the predominant usage of \eval in Base, through function \c{match.args}. In CRAN, it correlates with more exotic environments passed to \eval.

@@

Package \c{tidyselect} is the backend of several \c{dplyr} and \c{tidyr} functions to select columns. It offers a DSL that makes it possible to select a set of columns, over a range of columns, a complement, an intersection, an union, with a regular expression and so on. \eval is used to traverse selectors stored as slots of a dataframe.
